A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Additional Methods

### 2020-01-12 writeUp/supplementary_materials_methods.tex
added info on max age class

In the field, unfortunately, we do not know the number of mosquitoes recruited into the adult population, $A(a)$, nor the probability of capturing an individual at a given point in time, $p$. Instead, we model their product $\Psi=A(0)p$ (the population of adult female mosquitoes of age zero that are expected to be captured) probabilistically resulting in a model,
%
\begin{align}\label{eq:dissertation_negativeBinomial}
y(a) \sim \text{NB}(\Psi S(a),\kappa).
\end{align}
%
The resultant likelihood of a data series consisting of counts: $(y(a_1),y(a_2),...,y(a_R))$, is then calculated by assuming (conditional) independence of the observations,
%
\begin{equation}
\mathcal{L}(y(a_1),y(a_2),...,y(a_R)|S(.),\Psi,\kappa) = \prod\limits_{a=a_1}^{a_R} p(y(a)|S(a),\Psi,\kappa),
\end{equation}
%
where $p(y(a)|S(a),\Psi,\kappa)$ corresponds to the negative binomial probability mass function for a count of $y(a)$ mosquitoes aged $a$ as specified in eqn. (\ref{eq:dissertation_negativeBinomial}), and $R$ is the number of separate physiological age classes recorded.
%
Since we do not know $\Psi$, we must learn it from the data. One approach to estimate this parameter could be use the number of captures of nulliparous mosquitoes, $X(0)$ in place of $\Psi$. We prefer to allow for some uncertainty in this parameter and use the data to estimate its value. However, unfortunately, the negative binomial likelihood allows too much variation in this parameter (because the data are over-dispersed), and instead we specify a likelihood of the form,
%
\begin{align}
X(0) \sim \mathcal{N}(\Psi,\sqrt{\Psi}),
\end{align}
%
solely for the first data point $X(0)$. The above allows for some freedom in $\Psi$ whilst ensuring that the parameter's probability mass lies near enough to $X(0)$ to allow useful model estimates. $\Psi$ is set a uniform prior over the range $[0,\frac{3}{2}X(0)]$.

>>>>

In the field, unfortunately, we do not know the number of mosquitoes recruited into the adult population, $A(a)$, nor the probability of capturing an individual at a given point in time, $p$. Instead, we model their product $\Psi=A(0)p$ (the population of adult female mosquitoes of age zero that are expected to be captured) probabilistically resulting in a model,
%
\begin{align}\label{eq:dissertation_negativeBinomial}
y(a) \sim \text{NB}(\Psi S(a),\kappa).
\end{align}
%
The resultant likelihood of a data series consisting of counts: $(y(a_1),y(a_2),...,y(a_R))$, is then calculated by assuming (conditional) independence of the observations,
%
\begin{equation}
\mathcal{L}(y(a_1),y(a_2),...,y(a_R)|S(.),\Psi,\kappa) = \prod\limits_{a=a_1}^{a_R} p(y(a)|S(a),\Psi,\kappa),
\end{equation}
%
where $p(y(a)|S(a),\Psi,\kappa)$ corresponds to the negative binomial probability mass function for a count of $y(a)$ mosquitoes aged $a$ as specified in eqn. (\ref{eq:dissertation_negativeBinomial}), and $R$ is the number of separate physiological age classes that we include: for series with no censoring (see below for how censored data were handled), we include 14 separate age classes (from nulliparous to 13-parous). We recognise that the choice of a "maximum" age class is an approximation, since, in theory, we should include a contribution from each "0" observed until $\infty$. In reality, however, the limit was chosen to match the maximum number of cycles observed in the data, and the majority of series had far maxima that were less than 5. As such, we believe that our truncation should provide a good approximation to the infinite sum. 
%
Since we do not know $\Psi$, we must learn it from the data. One approach to estimate this parameter could be use the number of captures of nulliparous mosquitoes, $X(0)$ in place of $\Psi$. We prefer to allow for some uncertainty in this parameter and use the data to estimate its value. However, unfortunately, the negative binomial likelihood allows too much variation in this parameter (because the data are over-dispersed), and instead we specify a likelihood of the form,
%
\begin{align}
X(0) \sim \mathcal{N}(\Psi,\sqrt{\Psi}),
\end{align}
%
solely for the first data point $X(0)$. The above allows for some freedom in $\Psi$ whilst ensuring that the parameter's probability mass lies near enough to $X(0)$ to allow useful model estimates. $\Psi$ is set a uniform prior over the range $[0,\frac{3}{2}X(0)]$.
